---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# geneticMapRFiles

This repository stores data files used alongside the [geneticMapR](https://vegaalfaro.github.io/geneticMapR/) R package.

## Repository Purpose

This data-only repository hosts example and large-scale datasets used for development, testing, and demonstration of the `geneticMapR` package. These datasets are not included in the main CRAN package due to size limitations.

## Folder Structure

- `IDs`: Sample names and pedigree
- `phenotype`: phenotypic data 
- `vcf`: genotypic data, usually in vcf format.
- `R_data`: RDAs and other R objects and datasets.

## Usage

These files can be accessed in your R workflows by downloading them directly or by linking this GitHub repo in your pipeline.

## Related Package

See the main package: [`geneticMapR`](https://github.com/vegaalfaro/geneticMapR)

## License

MIT

## Files Description

### IDs

- `ID-Pedigrees-Biparentals-2025-01-22.xlsx`: This excel file contains sample-level metadata for individuals in a genetic study which corresponds to samples in a VCF file and downstream analysis.

- `ID.csv`: This csv file contains ID data for individuals in our genetic study which corresponds to samples in a VCF file and downstream analysis. The CSV format let us use it easily. It is the same information as sheet 2 of the file above.


### phenotype

- `phenotypes-binary.csv`: Phenotypic data of root shapes for an F~2~ mapping population of *Beta vulgaris*. Data includes unique identifiers and sample identifiers. Phenotype columns start at column I.


### R_data

The **filtered_geno_matrices_1629.RData** file contains the following R objects: 

- `hom_phased_geno_1629`: A genotype matrix of 2845 markers by 104 individuals. It keeps only homozygous markers (where parents are fixed for either 0 or 2 alleles). Filtered for missing data.
- `het_phased_geno_1629`: A genotype matrix of  of 12,155 markers and 104 individuals. It keeps homozygous and heterozygous markers.
- `het_phased_geno_1629_filt`: A filtered version of `het_phased_geno_1620` with 6482 markers and 104 individuals. Filtering parameters and procedures are explained in this [article](https://vegaalfaro.github.io/geneticMapR/articles/Recode.html).
- markers_homozygous: dimensions of `hom_phased_geno_1629`
- markers_heterozygous: dimensions of `het_phased_geno_1629_filt`

The **binned_geno_1629.RData** file contains the following R objects: 

- `geno.het.bin`: A genotype matrix of 2480 **binned** markers by 100 individuals. Includes heterozygous and homozygous markers.

- `geno.hom.bin`: A genotype matrix of 998 **binned** markers by 100 individuals. Includes homozygous markers only.

- `dendro`: a ggplot object showing a dendrogram used to visualize relationship among samples.

The **LODmats_1629-2025-03-12.RData** file contains the following R objects: 

- `LODmat.het`: Large matrix containing LOD scores for each marker pair. Heterozygous markers and Homozygous markers.

- `LODmat.hom`:Large matrix containing LOD scores for each marker pair. Homozygous markers.

The **ordered_markers.RData** file:

is a large list with nine elements. Each element corresponds to a chromosome (e.g., CHR1). Each one of those elements contains objects, the original genotype matrix, the ordered genotype matrix and two plots (the order plot and the haplotype frequency plot).

The **genotype_matrices_hmm_pop1-2025-03-12.RData**  file:

is a file that contains the reconstructed genotype after using an HMM model and the viterbi algorithm. 

The **genetic_maps_pop2-2025-03-12.RData**  file:

is a file that contains the genetic maps for the reference genome order and optimized order estimated according to this [script](https://vegaalfaro.github.io/geneticMapR/articles/Haplotype_and_Map.html).

### vcf

- `SNP_updated_IDs_sorted2.vcf.gz`: This is a lightly filtered version of the VCF file that contains genetic markers and individuals. The marker names have been processed for clarity. 


__Last updated:__
```{r, echo=FALSE}
Sys.Date()
```


